%TC: macro \marginfootnote [other]
%TC: envir SCfigure [] other
%TC: macrocount beginSCfigure [figure]
\documentclass[11pt,twoside]{report}
\usepackage{preamble}
\setcounter{chapter}{1}
\graphicspath{{../img/}}
\def\includebibliography{}

\externaldocument{introduction}
\externaldocument{morphometric-framework}
\externaldocument{morphometric-applications}
\externaldocument{resummation}
\externaldocument{aerosols}

\begin{document}
\chapter{Geometry and the liquid state}
\epigraph{It tells me that the Creator used the wrong kind of circles.}{Terry Pratchett, \emph{Pyramids} (1989).}
%\epigraph{It was not certain what significance the ceremony held... but the formality was no less sacred for it being unintelligible.}{Mervyn Peake, \emph{Titus Groan}, (1946).}
\label{chapter:background}

In this chapter we provide a self-contained account of the foundational frameworks underlying the two themes of this thesis: \emph{integral geometry} and \emph{liquid state theory}.
I expect the reader to have a background in statistical physics, so my account of liquid state theory is not intended to be exhaustive; for more in-depth treatments see the references herein.
By contrast, I do \emph{not} expect much familiarity with integral geometry.
Understanding the underlying mathematical detail of this field is not essential to follow the rest of the thesis, so I will focus on the key concepts and notation rather than detailed derivation.

This chapter was assembled from my notes on liquid state theory over the previous few years, which I wanted to organise in one place mostly for my own benefit.
As such, this chapter is a little long so I anticipate the expert reader will skim over it; to facilitate this I have placed the important results in boxes as a guide to the most relevant parts.

As many-body correlation functions are a central theme of the results chapters, I have emphasised correlation functions in section \ref{sec:liquid-structure} on liquid structure to the point where I have somewhat belaboured giving the explicit forms and normalisations of the various correlation functions.
Even though we will only use one particular hierarchy of correlation functions in the results chapters, I personally found it helpful to have these formulas in one place.
I have found myself frequently revisiting the transformations between the various hierarchies of correlation functions, so I include them in anticipation that someone repeating or extending this work may profit from having a kind of ``cheat sheet''.

\input{integral-geometry}
\input{liquid-state-theory}
\input{density-functional-theory}

\section{Summary}

\begin{SCtable}
  \begin{minipage}[b]{\linewidth}
    \centering
    \begin{tabular}{cc}
      \toprule
      \multicolumn{2}{c}{Dual hierarchies} \\
      \midrule
      correlation function & dual function \\
      \midrule
      distribution $g^{(n)}$ \eqref{eq:n-particle-distribution} &
      cluster $h^{(n)}$ \eqref{eq:cluster-correlation-functions} \\
      density moments $\langle \hat{\rho}_1 \cdots \hat{\rho}_n \rangle$ \eqref{eq:density-moments} &
      density-density $H^{(n)}$ \eqref{eq:density-density-correlations} \\
      Boltzmann weight $e^{-\beta U_n}$ &
      Ursell $W^{(n)}$ \eqref{eq:ursell-functions} \\
      \midrule
      \multicolumn{2}{c}{Other correlation functions} \\
      \midrule
      \multicolumn{2}{c}{$n$-particle density $\rho^{(n)}$ \eqref{eq:n-particle-density}} \\
      \multicolumn{2}{c}{static structure factor $S^{(n)}$ \eqref{eq:static-structure-factor}} \\
      \multicolumn{2}{c}{direct $c^{(n)}$ \eqref{eq:direct-correlations}} \\
      \bottomrule
    \end{tabular}
  \end{minipage}
  \caption{Summary of the various correlation functions in liquid state theory.}
  \label{table:correlation-functions}
\end{SCtable}

We have seen that integral geometry provides the formalism for describing sizes, and thus defines an \emph{ansatz} for strictly extensive variables.
Specifically, such variables must be a linear combination of the $(d+1)$ intrinsic volumes, which are the only meaningful size measures in $d$-dimensions in the sense of rigid-motion invariance, additivity and continuity.
We expect in general that thermodynamic potentials are extensive in the thermodynamic limit, so this formalism provides a good starting point for their description.
However, it is worth noting that extensive quantities could contain subleading terms which disappear in the thermodynamic limit; these would not be captured by an integral geometric \emph{ansatz}.

We introduced liquid state theory, with a particular focus on the various correlation functions which are summarised in Table \ref{table:correlation-functions}.
A key result for hard spheres was the highly accurate Carnahan-Starling equation of state \eqref{eq:cs-pressure}, which we will use throughout to treat the hard sphere liquid.
We saw how the intrinsic volumes could be applied to liquid state theory to construct accurate free energy functionals for hard particle systems.
We obtained the morphometric approach, an expansion of the chemical potential with the properties of sizes outlined above, as a special case of this theory.
In chapters \ref{chapter:morphometric-framework} and \ref{chapter:resummation} we will provide additional justifications of this approach, and we will apply it to predict the concentrations of local structures in chapter \ref{chapter:morphometric-applications}.

In the next chapter we will discuss the phenomenology of supercooled liquids and the glass transition, which will provide the key context for the treatment of the high density hard sphere liquid in subsequent chapters.

\ifdefined\includebibliography
  \newgeometry{margin=1in}
  \printbibliography
\fi

\end{document}
